---
title: "VAST Challenge 1"
output: html_document
---

#Loading and Installing Required Packages

```{r}
packages = c('ggplot2', 'gtable', 'grid', 'tuneR', 'audio', 'soundgen', 'warbleR')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

#Converting mp3 files of Actual Rose-Crested Blue Pipits recordings to wav files

```{r}
#Green_Song
#162563, 277952, 293914, 377874, 30397

realpipitgreensong1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-162563.mp3")
#writeWave(realpipitgreensong1, "realpipitgreensong1.wav", extensible=FALSE)

realpipitgreensong2 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-277952.mp3")
#writeWave(realpipitgreensong2, "realpipitgreensong2.wav", extensible=FALSE)

realpipitgreensong3 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-293914.mp3")
#writeWave(realpipitgreensong3, "realpipitgreensong3.wav", extensible=FALSE)

realpipitgreensong4 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-377874.mp3")
#writeWave(realpipitgreensong4, "realpipitgreensong4.wav", extensible=FALSE)

realpipitgreensong5 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-30397.mp3")
#writeWave(realpipitgreensong5, "realpipitgreensong5.wav", extensible=FALSE)

#Green_Call
#181907, 111775, 162566, 368492, 298739
realpipitgreencall1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-181907.mp3")
#writeWave(realpipitgreencall1, "realpipitgreencall1.wav", extensible=FALSE)

realpipitgreencall2 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-111775.mp3")
#writeWave(realpipitgreencall2, "realpipitgreencall2.wav", extensible=FALSE)

realpipitgreencall3 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-293916.mp3")
#writeWave(realpipitgreencall3, "realpipitgreencall3.wav", extensible=FALSE)

realpipitgreencall4 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-368492.mp3")
#writeWave(realpipitgreencall4, "realpipitgreencall4.wav", extensible=FALSE)

realpipitgreencall5 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-298739.mp3")
#writeWave(realpipitgreencall5, "realpipitgreencall5.wav", extensible=FALSE)

#Red_Song
#162564, 138985, 405548
realpipitredsong1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-162564.mp3")
#writeWave(realpipitredsong1, "realpipitredsong1.wav", extensible=FALSE)

realpipitredsong2 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-138985.mp3")
#writeWave(realpipitredsong2, "realpipitredsong2.wav", extensible=FALSE)

realpipitredsong3 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-405548.mp3")
#writeWave(realpipitredsong3, "realpipitredsong3.wav", extensible=FALSE)

#Red_call
#162567, 162569, 368493
realpipitredcall1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-162567.mp3")
#writeWave(realpipitredcall1, "realpipitredcall1.wav", extensible=FALSE)

realpipitredcall2 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-162569.mp3")
#writeWave(realpipitredcall2, "realpipitredcall2.wav", extensible=FALSE)

realpipitredcall3 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-368493.mp3")
#writeWave(realpipitredcall3, "realpipitredcall3.wav", extensible=FALSE)

#Purple_Song
#134557, 152971
realpipitpurplesong1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-134557.mp3")
#writeWave(realpipitpurplesong1, "realpipitpurplesong1.wav", extensible=FALSE)

realpipitpurplesong2 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-152971.mp3")
#writeWave(realpipitpurplesong2, "realpipitpurplesong2.wav", extensible=FALSE)

#Purple_Call
#405901
realpipitpurplecall1 <- readMP3("data/ALL BIRDS/Rose-Crested-Blue-Pipit-405901.mp3")
#writeWave(realpipitpurplecall1, "realpipitpurplecall1.wav", extensible=FALSE)

```

#Commands to Play Converted Wav Files for test recordings

```{r}
play(load.wave("realpipitgreensong1.wav"))
```

#Plotting Actual Pipit Recordings (Green Song)

```{r}
spectro(realpipitgreensong1,cont=TRUE, scale=FALSE, main = "Green Song 1: 162563")

spectro(realpipitgreensong2,cont=TRUE, scale=FALSE, main = "Green Song 2: 277952")

spectro(realpipitgreensong3,cont=TRUE, scale=FALSE, main = "Green Song 3: 293914")

spectro(realpipitgreensong4,cont=TRUE, scale=FALSE, main = "Green Song 4: 377874")

spectro(realpipitgreensong5,cont=TRUE, scale=FALSE, main = "Green Song 5: 30397")

spectro(realpipitgreensong1,osc=TRUE)

spectro(realpipitgreensong2,osc=TRUE)

spectro(realpipitgreensong3,osc=TRUE)

spectro(realpipitgreensong4,osc=TRUE)

spectro(realpipitgreensong5,osc=TRUE)
```

#Plotting Actual Pipit Recordings (Green Call)

```{r}
spectro(realpipitgreencall1,cont=TRUE, scale=FALSE, main = "Green Call 1: 181907")

spectro(realpipitgreencall2,cont=TRUE, scale=FALSE, main = "Green Call 2: 111775")

spectro(realpipitgreencall3,cont=TRUE, scale=FALSE, main = "Green Call 3: 293916")

spectro(realpipitgreencall4,cont=TRUE, scale=FALSE, main = "Green Call 4: 368492")

spectro(realpipitgreencall5,cont=TRUE, scale=FALSE, main = "Green Call 5: 298739")

spectro(realpipitgreencall1,osc=TRUE)

spectro(realpipitgreencall2,osc=TRUE)

spectro(realpipitgreencall3,osc=TRUE)

spectro(realpipitgreencall4,osc=TRUE)

spectro(realpipitgreencall5,osc=TRUE)
```

#Plotting Actual Pipit Recordings (Red Song)

```{r}
spectro(realpipitredsong1,cont=TRUE, scale=FALSE, main = "Red Song 1: 162564")

spectro(realpipitredsong2,cont=TRUE, scale=FALSE, main = "Red Song 2: 138985")

spectro(realpipitredsong3,cont=TRUE, scale=FALSE, main = "Red Song 3: 405548")

spectro(realpipitredsong1,osc=TRUE)

spectro(realpipitredsong2,osc=TRUE)

spectro(realpipitredsong3,osc=TRUE)

```

#Plotting Actual Pipit Recordings (Red Call)

```{r}
spectro(realpipitredcall1,cont=TRUE, scale=FALSE, main = "Red Call 1: 162567")

spectro(realpipitredcall2,cont=TRUE, scale=FALSE, main = "Red Call 2: 162569")

spectro(realpipitredcall3,cont=TRUE, scale=FALSE, main = "Red Call 3: 368493")

spectro(realpipitredcall1,osc=TRUE)

spectro(realpipitredcall2,osc=TRUE)

spectro(realpipitredcall3,osc=TRUE)

```

#Plotting Actual Pipit Recordings (Purple Song)

```{r}
spectro(realpipitpurplesong1,cont=TRUE, scale=FALSE, main = "Purple Song 1: 134557")

spectro(realpipitpurplesong2,cont=TRUE, scale=FALSE, main = "Purple Song 2: 152971")

spectro(realpipitpurplesong1,osc=TRUE)

spectro(realpipitpurplesong2,osc=TRUE)
```

#Plotting Actual Pipit Recordings (Purple Call)

```{r}
spectro(realpipitpurplecall1,cont=TRUE, scale=FALSE, main = "Purple Call 1: 405901")


spectro(realpipitpurplecall1,osc=TRUE)
```
#Converting mp3 files of "Test Birds from Kasios" to wav files

```{r}
#Green
#1, 6, 11, 15
testpipitgreen1 <- readMP3("data/Test Birds from Kasios/1.mp3")
#writeWave(testpipitgreen1, "testpipitgreen1.wav", extensible=FALSE)

testpipitgreen6 <- readMP3("data/Test Birds from Kasios/6.mp3")
#writeWave(testpipitgreen6, "testpipitgreen6.wav", extensible=FALSE)

testpipitgreen11 <- readMP3("data/Test Birds from Kasios/11.mp3")
#writeWave(testpipitgreen11, "testpipitgreen11.wav", extensible=FALSE)

testpipitgreen15 <- readMP3("data/Test Birds from Kasios/15.mp3")
#writeWave(testpipitgreen15, "testpipitgreen15.wav", extensible=FALSE)

#Red
#2, 3, 4, 7, 9, 13, 14
testpipitred2 <- readMP3("data/Test Birds from Kasios/2.mp3")
#writeWave(testpipitred2, "testpipitred2.wav", extensible=FALSE)

testpipitred3 <- readMP3("data/Test Birds from Kasios/3.mp3")
#writeWave(testpipitred3, "testpipitred3.wav", extensible=FALSE)

testpipitred4 <- readMP3("data/Test Birds from Kasios/4.mp3")
#writeWave(testpipitred4, "testpipitred4.wav", extensible=FALSE)

testpipitred7 <- readMP3("data/Test Birds from Kasios/7.mp3")
#writeWave(testpipitred7, "testpipitred7.wav", extensible=FALSE)

testpipitred9 <- readMP3("data/Test Birds from Kasios/9.mp3")
#writeWave(testpipitred9, "testpipitred9.wav", extensible=FALSE)

testpipitred13 <- readMP3("data/Test Birds from Kasios/13.mp3")
#writeWave(testpipitred13, "testpipitred13.wav", extensible=FALSE)

testpipitred14 <- readMP3("data/Test Birds from Kasios/14.mp3")
#writeWave(testpipitred14, "testpipitred14.wav", extensible=FALSE)

#Purple
#5, 8, 10, 12

testpipitpurple5 <- readMP3("data/Test Birds from Kasios/5.mp3")
#writeWave(testpipitpurple5, "testpipitpurple5.wav", extensible=FALSE)

testpipitpurple8 <- readMP3("data/Test Birds from Kasios/8.mp3")
#writeWave(testpipitpurple8, "testpipitpurple8.wav", extensible=FALSE)

testpipitpurple10 <- readMP3("data/Test Birds from Kasios/10.mp3")
#writeWave(testpipitpurple10, "testpipitpurple10.wav", extensible=FALSE)

testpipitpurple12 <- readMP3("data/Test Birds from Kasios/12.mp3")
#writeWave(testpipitpurple12, "testpipitpurple12.wav", extensible=FALSE)

```
#Commands to Play Converted Wav Files for test recordings

```{r}
play(load.wave("testpipitgreen1.wav"))
```

#Plotting Test Pipit Recordings (Green)

```{r}
spectro(testpipitgreen1,cont=TRUE, scale=FALSE, main = "Test Green 1")

spectro(testpipitgreen6,cont=TRUE, scale=FALSE, main = "Test Green 6")

spectro(testpipitgreen11,cont=TRUE, scale=FALSE, main = "Test Green 11")

spectro(testpipitgreen15,cont=TRUE, scale=FALSE, main = "Test Green 15")


spectro(testpipitgreen1,osc=TRUE)

spectro(testpipitgreen6,osc=TRUE)

spectro(testpipitgreen11,osc=TRUE)

spectro(testpipitgreen15,osc=TRUE)
```


#Plotting Test Pipit Recordings (Red)

```{r}
spectro(testpipitred2,cont=TRUE, scale=FALSE, main = "Test Red 2")

spectro(testpipitred3,cont=TRUE, scale=FALSE, main = "Test Red 3")

spectro(testpipitred4,cont=TRUE, scale=FALSE, main = "Test Red 4")

spectro(testpipitred7,cont=TRUE, scale=FALSE, main = "Test Red 7")

spectro(testpipitred9,cont=TRUE, scale=FALSE, main = "Test Red 9")

spectro(testpipitred13,cont=TRUE, scale=FALSE, main = "Test Red 13")

spectro(testpipitred14,cont=TRUE, scale=FALSE, main = "Test Red 14")


spectro(testpipitred2,osc=TRUE)

spectro(testpipitred3,osc=TRUE)

spectro(testpipitred4,osc=TRUE)

spectro(testpipitred7,osc=TRUE)

spectro(testpipitred9,osc=TRUE)

spectro(testpipitred13,osc=TRUE)

spectro(testpipitred14,osc=TRUE)
```

#Plotting Test Pipit Recordings (Purple)

```{r}
spectro(testpipitpurple5,cont=TRUE, scale=FALSE, main = "Test Purple 5")

spectro(testpipitpurple8,cont=TRUE, scale=FALSE, main = "Test Purple 8")

spectro(testpipitpurple10,cont=TRUE, scale=FALSE, main = "Test Purple 10")

spectro(testpipitpurple12,cont=TRUE, scale=FALSE, main = "Test Purple 12")


z <- readWave("testpipitpurple8.wav", from = 15, to = 18, units = "seconds")
spectro(z)
spectro(testpipitpurple8,cont=TRUE, scale=FALSE)

spectro(testpipitpurple5,osc=TRUE)

spectro(testpipitpurple8,osc=TRUE)

spectro(testpipitpurple10,osc=TRUE)

spectro(testpipitpurple12,osc=TRUE)
```



